### 2017-08-24 sec-conclusion.tex
Whitespace changes and minor edits.

\section{Conclusions and Future Work}
% summarize the main contributions of the paper; summarize the experimental
% results; argue again for our novelty; discuss future directions
As a tool to help instructors and education researchers better understand
the behavior of MOOC students, we proposed a two-layer hidden Markov model
to automatically extract student activity patterns in the form of behavior
state-transition graphs from large amounts of MOOC log data. This model is
different from existing methods in that it treats behavior patterns as a
sequence of \emph{latent states}, rather than assigning these states in a
rule-based manner. It captures the variable behaviors of students over time
and allows analysis at different levels of granularity.

>>>>

\section{Conclusions and Future Work}
As a tool to help instructors and education researchers better understand
the behavior of MOOC students, we proposed a two-layer hidden Markov model
to automatically extract student activity patterns in the form of behavior
state-transition graphs from large amounts of MOOC log data. This model is
different from existing methods in that it treats behavior patterns as a
sequence of \emph{latent states}, rather than assigning these states in a
rule-based manner. It captures the variable behaviors of students over time
and allows analysis at different levels of granularity.

### 2017-08-24 sec-intro.tex
Whitespace changes and minor edits.

Thus, what we propose is a \emph{latent variable approach} for mining student behavior
patterns that is \emph{probabilistic} for inference and
does \emph{not} force assumptions about time resolutions, making it
\emph{flexible to model state changes over different time resolutions} more
easily.  More specifically, we propose a novel two-layer hidden Markov
model (2L-HMM) to discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences.  Evaluation results on a MOOC data set on Coursera demonstrate
that the 2L-HMM can effectively discover a variety of interesting interpretable
student behavior patterns at different levels of resolution, many of which
are beyond what existing approaches can discover. We show that the patterns
uncovered by the 2L-HMM capture meaningful behavior by quantitatively
showing that features extracted from a trained 2L-HMM correlate with
learning outcomes. Since our proposed methods are
unsupervised, they can potentially be applied to any MOOC data without
requiring manual annotation effort at the level of sequences. Instead,
instructors are empowered to use the latent patterns that the 2L-HMM can
discover from raw data to extract knowledge about the behaviors his/her
students exhibit in the MOOC.

>>>>

Thus, what we propose is a \emph{latent variable approach} for mining
student behavior patterns that is \emph{probabilistic} for inference and
does \emph{not} force assumptions about time resolutions, making it
\emph{flexible to model state changes over different time resolutions} more
easily.  More specifically, we propose a novel two-layer hidden Markov
model (2L-HMM) to discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences.  Evaluation results on a MOOC data set on Coursera demonstrate
that the 2L-HMM can effectively discover a variety of interesting
interpretable student behavior patterns at different levels of resolution,
many of which are beyond what existing approaches can discover. We show
that the patterns uncovered by the 2L-HMM capture meaningful behavior by
quantitatively showing that features extracted from a trained 2L-HMM
correlate with learning outcomes. Since our proposed methods are
unsupervised, they can potentially be applied to any MOOC data without
requiring manual annotation effort at the level of sequences, thus
empowering instructors to use the latent patterns discovered by the 2L-HMM
to further extract knowledge about the behaviors his/her students exhibit
in the MOOC.

### 2017-08-24 sec-model.tex
Whitespace changes and minor edits.

Given a MOOC log, we can define a set $\mathbf{A}$ of actions that a
student can take at any given time. For example, an action $a \in
\mathbf{A}$ might be ``viewing lecture,'' ``taking quiz,'' or ``viewing
forum.'' For each student in the course $\ell \in \mathbf{L}$, we then
extract a list of action sequences $\mathbf{O}_\ell$ that he or she
produced as observed in the log, where each sequence $\mathbf{o} \in
\mathbf{O}_\ell$ is itself a list of actions $(a_1, a_2, \ldots, a_T)$ with
each $a_i \in \mathbf{A}$.  Each sequence can be divided flexibly; in this
paper, we chose to denote the end of a sequence as occurring when no further
actions occur within a 10-hour window of time (and thus our sequences
roughly correspond to one day's worth of activity). This decision reflects
our desire to uncover latent state transition behavior at the granularity
of day-to-day behavior. A different segmentation strategy could be used to
uncover hour-by-hour behavior or week-by-week behavior, and this depends
entirely on the desired time resolution one wishes to be exhibited in the
transitions between the latent states. Different segmentation strategies
will result in different underlying training data, and thus different
meanings behind the patterns that the 2L-HMM will extract. The flexibility of
using different segmentation strategies is intentional as it allows a user to
flexibly adjust the segmentation so as to obtain patterns at different granularity levels as needed.

>>>>

Given a MOOC log, we can define a set $\mathbf{A}$ of actions that a
student can take at any given time. For example, an action $a \in
\mathbf{A}$ might be ``viewing lecture,'' ``taking quiz,'' or ``viewing
forum.'' For each student in the course $\ell \in \mathbf{L}$, we then
extract a list of action sequences $\mathbf{O}_\ell$ that he or she
produced as observed in the log, where each sequence $\mathbf{o} \in
\mathbf{O}_\ell$ is itself a list of actions $(a_1, a_2, \ldots, a_T)$ with
each $a_i \in \mathbf{A}$.  Each sequence can be divided flexibly; in this
paper, we chose to denote the end of a sequence as occurring when no further
actions occur within a 10-hour window of time (and thus our sequences
roughly correspond to one day's worth of activity). This decision reflects
our desire to uncover latent state transition behavior at the granularity
of day-to-day behavior. A different segmentation strategy could be used to
uncover hour-by-hour behavior or week-by-week behavior, and this depends
entirely on the desired time resolution one wishes to be exhibited in the
transitions between the latent states. Different segmentation strategies
will result in different underlying training data, and thus different
meanings behind the patterns that the 2L-HMM will extract. The flexibility
of using different segmentation strategies is intentional as it allows a
user to adjust the segmentation as needed so as to obtain patterns at
different granularity levels.
